%
% personal commentary:
%        DRAFT DRAFT DRAFT
%        - GNGUYEN
%
\chapter{Error Model}
\label{chap:error_model}

This chapter describes the implementation and configuration of error
models, which introduces packet losses into a simulation. 

In addition to the basic class ErrorModel described in details below, 
  there are several other types of error modules not being 
  completely documented yet,
  which include:
  \begin{itemize}
    \item SRMErrorModel, PGMErrorModel: 
      error model for SRM and PGM.
    \item ErrorModel/Trace: 
      error model that reads a loss trace (instead of a math/computed model)
    \item MrouteErrorModel:
      error model for multicast routing, now inherits from trace.
    \item ErrorModel/Periodic:
      models periodic packet drops (drop every nth packet we see).
      This model can be conveniently combined with a flow-based classifier
      to achieve drops in particular flows
    \item SelectErrorModel:
       for Selective packet drop.
    \item ErrorModel/TwoState:
      Two-State:  error-free and error
    \item ErrorModel/TwoStateMarkov, ErrorModel/Expo, ErrorModel/Empirical:
      inerit from ErrorModel/TwoState.
    \item ErrorModel/List:
      specify a list of packets/bytes to drop,
      which could be in any order.
  \end{itemize}

Their definitions can be found in \nsf{queue/errmodel.\{cc, h\}} and
 \nsf{tcl/lib/ns-errmodel.tcl, ns-default.tcl}.

\section{Implementation}

The procedures and functions described in this section can be found in
\nsf{errmodel.\{cc, h\}}.

Error model simulates link-level errors or loss by either marking the
packet's error flag or dumping the packet to a drop target.  In
simulations, errors can be generated from a simple model such as the
packet error rate, or from more complicated statistical and empirical models.
To support a wide variety of models, the unit of error can be specified
in term of packet, bits, or time-based.

The \code{ErrorModel} class is derived from the \code{Connector} base
class.  As the result, it inherits some methods for hooking up objects
such as \code{target} and \code{drop-target}.  If the drop target
exists, it will received corrupted packets from \code{ErrorModel}.
Otherwise, \code{ErrorModel} just marks the \code{error_} flag of the
packet's common header, thereby, allowing agents to handle the loss.
The \code{ErrorModel} also defines additional Tcl method \code{unit} to
specify the unit of error and \code{ranvar} to specify the random
variable for generating errors.  If not specified, the unit of error
will be in packets, and the random variable will be uniform distributed
from 0 to 1.  Below is a simple example of creating an error model with
the packet error rate of 1 percent (0.01):
\begin{program}
        # create a loss_module and set its packet error rate to 1 percent
        set loss_module [new ErrorModel]
        \$loss_module set rate_ 0.01

        # {\cf optional:  set the unit and random variable}
        \$loss_module unit pkt            \; error unit: packets (the default);
        \$loss_module ranvar [new RandomVariable/Uniform]

        # {\cf set target for dropped packets}
        \$loss_module drop-target [new Agent/Null]
\end{program}

In C++, the \code{ErrorModel} contains both the mechanism and policy for
dropping packets.  The packet dropping mechanism is handled by the
\code{recv} method, and packet corrupting policy is handled by the
\code{corrupt} method.
\begin{program}
        enum ErrorUnit \{ EU_PKT=0, EU_BIT, EU_TIME \};

        class ErrorModel : public Connector \{
        public:
                ErrorModel();
                void recv(Packet*, Handler*);
                virtual int corrupt(Packet*);
                inline double rate() \{ return rate_; \}
        protected:
                int command(int argc, const char*const* argv);
                ErrorUnit eu_;          \* error unit in pkt, bit, or time */
                RandomVariable* ranvar_;
                double rate_;
        \};
\end{program}
The \code{ErrorModel} only implements a simple policy based on a single
error rate, either in packets of bits.  More sophisticated dropping
policy can be implemented in C++ by deriving from \code{ErrorModel} and
redefining its \code{corrupt} method.

%{\bf To be completed.}

\section{Configuration}

The previous section talked about error model, in this section we
discuss how to use error models in ns over either wired networks or
wireless ones. 

To use an error model for wired networks, at first it has to be
inserted into a SimpleLink object. Because a SimpleLink is a composite
object (Chapter \ref{chap:links}), an error model can be inserted to
many places. Currently we provide the following methods to insert an
error module into three different places.

\begin{itemize}
\item Insert an error module in a SimpleLink BEFORE the queue module. 
  This is provided by the following two OTcl methods: 

  \begin{alist}
    SimpleLink::errormodule args & When an error model is given
    as a parameter, it inserts the error module into the simple link,
    right after the queue module, and set the drop-target of the error
    model to be the drop trace object of the simple link. Note that
    this requires the following configuration order:
	\code{ns namtrace-all}
  followed by link configurations, followed by error
    model insertion. When no argument is given, it returns the current
    error model in the link, if there's any. This method is defined in
    \ns/tcl/lib/ns-link.tcl \\

    Simulator::lossmodel \tup{em} \tup{src} \tup{dst} & Call
    SimpleLink::errormodule to insert the given error module into the
    simple link (src, dst). It's simply a wrapper for the above
    method. This method is defined in \ns/tcl/lib/ns-lib.tcl. 
  \end{alist}
  
\item Insert an error module in a SimpleLink AFTER the queue but
  BEFORE the delay link.
  This is provided by the following two methods:

  \begin{alist}
    SimpleLink::insert-linkloss args & This method's behavior
    is identical to that of \code{SimpleLink::errormodule}, except
    that it inserts an error module immediately after the queue
    object. It's defined in \ns/tcl/lib/ns-link.tcl \\

    Simulator::link-lossmodel \tup{em} \tup{src} \tup{dst} &
    This is a wrapper for \code{SimpleLink::insert-linkloss}. It's
    defined in \ns/tcl/lib/ns-lib.tcl 
  \end{alist}
  
  The nam traces generated by error models inserted using these two
  methods do not require special treatment and can be visualized using
  an older version of nam. 

\item Insert an error module in a Link AFTER the delay link module. 
  This can be done by \code{Link::install-error}. 
  Currently this API doesn't produce any trace. It only serves as a
  placeholder for possible future extensions. 
\end{itemize}

To add an error model over wireless networks, each node can insert 
a given statistical error model either over outgoing or incoming wireless channels.
Precisely, the instanciated error model would be stuck between mac and netif modules 
depicted in Figure \ref{fig:mobilenode-dsr}. For the outgoing link, the error module
would be pointed by downtarget\_ of the above mac module while for the incoming link 
it would be linked by uptaget\_ pointer of the below netif module. And in each case 
the target\_ of the error module points to either the netif or the mac respectively.  
The difference of placing over the two different locations is that the outgoing causes
all the receivers to receive the packet suffering the same degree of errors since the error is
determined before wireless channel module copies the packet. On the other
hand, the incoming error module lets each receiver get the packet corrupted with different
degree of error since the error is independently computed in each error module.

The insertion into the wireless protocol stack can be done by calling
node-config command explained in Section~\ref{sec:node:nodeconfig} with
the two options IncomingErrrProc and OutgoingErrProc. We can use these two options 
at the same time or each one separately.  The argument of the two option is the
name of the global procedure which creates an error model object,
feeds it with necessary initial values appropriate for the new error
module, and finally returns the object pointer. The following shows a
simple TCL example script to add an error module into the wireless
protocol stack.
\begin{program}
	$ns node-config -IncomingErrProc UniformErr -OutgoingErrProc UniformErr

	proc UniformErr {} {
		set err [new ErrorModel]
		$err unit packet
		return $err
	}
\end{program}

\section{Multi-state error model}

Contributed by Jianping Pan (jpan@bbcr.uwaterloo.ca).

The multi-state error model implements time-based error state
transitions. Transitions to the next  
error state occur at the end of the duration of the current state. The
next error state is then selected using the 
transition state matrix.

To create a multi-state error model, the following parameters should
be supplied (as defined in \ns/tcl/lib/ns-errmodel.tcl): 
\begin{itemize}
\item {\tt states}: an array of states (error models).
\item {\tt periods}: an array of state durations.
\item {\tt trans}: the transition state model matrix.
\item {\tt transunit}: one of {\tt [pkt|byte|time]}.
\item {\tt sttype}: type of state transitions to use: either {\tt
    time} or {\tt pkt}.
\item {\tt nstates}: number of states.
\item {\tt start}: the start state.
\end{itemize}

Here is a simple example script to create a multi-state error model:
\begin{program}
        set tmp [new ErrorModel/Uniform 0 pkt]
        set tmp1 [new ErrorModel/Uniform .9 pkt]
        set tmp2 [new ErrorModel/Uniform .5 pkt]

        # {\cf Array of states (error models)}
        set m_states [list $tmp $tmp1 $tmp2]
        # {\cf Durations for each of the states, tmp, tmp1 and tmp2, respectively}
        set m_periods [list 0 .0075 .00375]
        # {\cf Transition state model matrix}
        set m_transmx \{ \{0.95 0.05 0\}
          \{0   0   1\}
          \{1   0   0\} \}
        set m_trunit pkt
        # {\cf Use time-based transition}
        set m_sttype time
        set m_nstates 3
        set m_nstart [lindex $m_states 0]

        set em [new ErrorModel/MultiState $m_states $m_periods $m_transmx \\
                $m_trunit $m_sttype $m_nstates $m_nstart]
\end{program} %$

\section{Commands at a glance}
\label{sec:errmodelcommand}

The following is a list of error-model related commands commonly used in
simulation scripts:

\begin{program}
set em [new ErrorModel]
$em unit pkt
$em set rate_ 0.02
$em ranvar [new RandomVariable/Uniform]
$em drop-target [new Agent/Null]
\end{program}

This is a simple example of how to create and configure an error model.
The commands to place the error-model in a simple link will be shown next.

\begin{flushleft}
\code{$simplelink errormodule  <args>}\\
This commands inserts the error-model before the queue object in simple link.
However in this case the error-model's drop-target points to the link's
\code{drophead_} element.


\code{$ns_ lossmodel <em> <src> <dst>}\\
This command places the error-model before the queue in a simplelink defined
by the <src> and <dst> nodes. This is basically a wrapper for the above method.


\code{$simplelink insert-linkloss <args>}\\
This inserts a loss-module after the queue, but right before the delay \code{link_}
element in the simple link. This is because nam can visualize a packet drop
only if the packet is on the link or in the queue. The error-module's
drop-target points to the link's \code{drophead_} element.


\code{$ns_ link-lossmodel <em> <src> <dst>}\\
This too is a wrapper method for insert-linkloss method described above.
That is this inserts the error-module right after the queue element in a
simple link (src-dst).

\end{flushleft}


% Local Variables:
% TeX-master: "everything"
% LocalWords:
% End
