
\chapter{Undocumented Facilities}

Ns is often growing to include new protocols.
Unfortunately the documention doesn't grow quite as often.
This section lists what remains to be documented,
  or what needs to be improved.

(The documentation is in the doc subdirectory of the ns source code
  if you want to add to it. :-)

\begin{description}
\item[Interface to the Interpreter]

        \begin{itemize}
        \item nothing currently
        \end{itemize}

\item[Simulator Basics]

        \begin{itemize}
        \item LANs need to be updated for new wired/wireless support
                (Yuri updated this?)
        \item wireless support needs to be added
                (done)
        \item should explicitly list queueing options in the queue mgt chapter?
        \end{itemize}

\item[Support]

        \begin{itemize}
        \item should pick a single list mgt package and document it
        \item should document the trace-post-processing utilities in bin
        \end{itemize}

\item[Routing]

  \begin{itemize}
  \item The usage and design of link state and MPLS routing modules
    are not documented at all. (Note: link state and MPLS appeared only in
    daily snapshots and releases after 09/14/2000.)
  \item need to document hierarchical routing/addressing
    (Padma has done)
  \item need a chapter on supported ad-hoc routing protocols
  \end{itemize}
  
\item[Queueing]

        \begin{itemize}
        \item CBQ needs documentation (can maybe build off of 
                \url{ftp://ftp.ee.lbl.gov/papers/cbqsims.ps.Z}?)
        \end{itemize}

\item[Transport]

        \begin{itemize}
        \item need to document MFTP
        \item need to document RTP (session-rtp.cc, etc.)
        \item need to document multicast building blocks
        \item should repair and document snoop and tcp-int
        \end{itemize}

\item[Traffic and scenarios] (new section)

        \begin{itemize}
        \item should add a description of how to drive the simulator
                from traces
        \item should add discussion of the scenario generator
        \item should add discussion of http traffic sources
        \end{itemize}

\item[Application]

        \begin{itemize}
        \item is the non-Haobo http stuff documented? no.
        \end{itemize}

\item[Scale]

        \begin{itemize}
        \item should add disucssion of mixed mode (pending)
        \end{itemize}

\item[Emulation]

        \begin{itemize}
        \item nothing currently
        \end{itemize}

\item[Other]

        \begin{itemize}
        \item should document admission control policies?
        \item should add a validation chapter and snarf
                up the contents of ns-tests.html
        \item should snarf up Marc Greis' tutorial
                rather than just referring to it?
        \end{itemize}

\end{description}

