\newcommand{\simname}{\em ns-2}
\newcommand{\codelocation}{\tt mash.cs.berkeley.edu}
\section Introduction
This manual is intended to be a general reference document
for people wishing to use or develop modules for the \simname
simulator.
The first section provides an overview of the development environment
and how to be sure your changes are properly integrated into the
source tree.
The second section details the architecture of \simname,
including how objects relate to one another, plus some examples.
The first two sections are intended to be most useful for people
developing new modules for \simname.
The next section attempts to document some of the more confusing
or complicated interactions between the various support libraries
needed by \simname.
The final section focuses on how to perform simulations and present
simulation results.

\subsection The environment

The sources for \simname are located on the machine \codelocation,
under control of CVS, a source code control system built on top of RCS.
The machine is presently configured to only allow authenticated
remote login using Kerberos.
In particular, encrypted remote login sessions using {\tt telnet -x} is
required.

There are several mailing lists related to the project:
\begin{itemize}
	\item[{\sf ns-developers@mash.cs.berkeley.edu] - current developers
	\item[{\sf ns-announce@mash.cs.berkeley.edu] - announcements of new releases
	\item[{\sf ns-users@mash.cs.berkeley.edu] - user community discussions
\end{itemize}
To subscribe to the lists, send mail to {\tt majordomo@mash.cs.berkeley.edu}.
Send just the single line {\tt help} in your message if you are unfamiliar
with the operation of the majordomo subscription service.

There are also several web pages of interest:
\begin{itemize}
	\item[{\sf http://www-mash.cs.berkeley.edu/ns}] - status
	\item[{\sf http://netweb.usc.edu/vint}] - description of VINT
	\item[{\sf http://www-nrg.ee.lbl.gov/ns}] - version 1 of {\tt ns}
\end{itemize}

Here is a (nonexhaustive) list of people involved in the project:
\begin{itemize}
	\item USC/ISI
	\begin{itemize}
	\item[estrin@usc.edu] - Deborah Estrin
	\item[dante@valhalla.internet-care.com] - Dante De Lucia
	\item[kannan@catarina.usc.edu] - Kannan Varadhan
	\item[ahelmy@catarina.usc.edu] - Ahmed A-G Helmy
	\item[wlee@isi.edu] - WeeSan Lee
	\item[daniel@isi.edu] - Danial Zappala
	\item[mjh@isi.edu] - Mark Handley
	\end{itemize}
	\item Xerox PARC
	\begin{itemize}
	\item[shenker@parc.xerox.com] - Scott Shenker
	\item[breslau@parc.xerox.com] - Lee Breslau
	\item[bajaj@parc.xerox.com] - Sandeep Bajaj
	\end{itemize}
	\item LBNL
	\begin{itemize}
	\item[floyd@ee.lbl.gov] - Sally Floyd
	\item[kfall@ee.lbl.gov] - Kevin Fall
	\item[van@ee.lbl.gov] - Van Jacobson
	\end{itemize}
	\item UCB
	\begin{itemize}
	\item[mccanne@eecs.berkeley.edu] - Steven McCanne
	\item[elan@mercenary.cs.berkeley.edu] - Elan Amir
	\item[tomh@kayak.cs.berkeley.edu] - Tom Henderson
	\end{itemize}
\end{itemize}


The normal procedure is to first set up a CVS environment (see below)
before proceeding.


\subsection CVS Basics
\subsection Incorporating changes
